### 2013-01-11 Docs/Thesis/background.tex
Take care of revisions from mtf

switcher :: Behavior a -> Event (Behavior a) -> Behavior a
switcher b [] t = b t
switcher b [(to, bo)] t = if t < to
                          then b t
                          else bo t
\end{code}
\hrule
\caption{An obvious, yet inefficient and problematic, implementation of
Classic FRP.}
\label{figure:classic_frp_simple_implementation}
\end{figure}

@@

An alternative approach to FRP was first proposed in work on
Fruit~\cite{Courtney2001-1}, a library for declarative specification of GUIs.
This library utilized the abstraction of Arrows~\cite{Hughes2000} to structure
{\em signal functions}. Arrows are abstract type constructors with input and
output type parameters, together with a set of routing combinators (see Figure~\ref{figure:arrow_combinators}).
Signal functions form an arrow. Signal functions are the first-class abstraction
in this approach. They represent time-dependent and reactive transformers of
events and signals, which are themselves not first class, since such values
cannot be directly manipulated by the programmer. This approach has two
motivations: it increases modularity since both the input and output of signal
functions may be transformed (as opposed to signals or events which may only be
transformed in their output) and it avoids a large class of time and space leaks
which have emerged when implementing FRP with first-class signals and events.

@@

A recent PhD thesis~\cite{Sculthorpe2011} introduced N-Ary FRP, a technique for
typing Arrowized FRP systems using dependent types. The bulk of the work
consisted in using the dependent type system to prove the correctness of the FRP
system discussed. This work introduced a signal vectors, a typing construct
which permits the distinction of signal and event types at the level of the FRP
system, instead of making events merely a special type of signal.

>>>>

switcher :: Behavior a -> Event (Behavior a) -> Behavior a
switcher b [] t = b t
switcher b ((to, bo):_) t = if t < to
                            then b t
                            else bo t
\end{code}
\hrule
\caption{An obvious, yet inefficient and problematic, implementation of
Classic FRP.}
\label{figure:classic_frp_simple_implementation}
\end{figure}

@@

An alternative approach to FRP was first proposed in work on
Fruit~\cite{Courtney2001-1}, a library for declarative specification of GUIs.
This library utilized the concept of Arrows~\cite{Hughes2000} as an abstraction
for {\em signal functions}. Arrows are abstract type constructors with input and
output type parameters, together with a set of routing combinators (see Figure~\ref{figure:arrow_combinators}).
Signal functions form an arrow. Signal functions represent time-dependent and
reactive transformers of events and signals. Signals and events
cannot be directly manipulated by the programmer. This approach has two
motivations: it increases modularity since both the input and output of signal
functions may be transformed (as opposed to signals or events which may only be
transformed in their output) and it avoids a large class of time and space leaks
which have emerged when implementing FRP with first-class signals and events.

@@

A recent PhD thesis~\cite{Sculthorpe2011} introduced N-Ary FRP, a technique for
typing Arrowized FRP systems using dependent types. The bulk of the work
consisted in using the dependent type system to prove the correctness of the FRP
system discussed. This work introduced signal vectors, a typing construct
which permits the distinction of signal and event types at the level of the FRP
system, instead of making events merely a special type of signal.

### 2013-01-11 Docs/Thesis/comparison.tex
Thesis pdf update
Take care of revisions from mtf

In both Yampa and TimeFlies, this algorithm is implemented by using an accumulator,
which applies functions carried by event occurrences to a state value. It stores
the new state and outputs new event occurrences based on the output of the function.

The state stored is the table (implemented using the {\tt Data.Map} module in the
Haskell {\tt containers} library) and the output event occurrences carry messages
to be sent to the switch. The events are sourced internally to the learning
switch signal function (using the {\tt after} combinator in TimeFlies and the {\tt repeatedly}
function in Yampa) to delete expired rules, and externally, by transforming messages
from the switch to the controller into accumulator functions.

@@

\begin{figure}
\epsfig{file=graph.eps,angle=-90,width=\linewidth}
\hrule
\caption{Comparison of timeflies vs. Yampa implementing an OpenFlow learning switch controller.}
\label{fig:timeflies-yampa-comparison}
\end{figure}

>>>>

In both Yampa and TimeFlies, this algorithm is implemented by using an accumulator,
which applies functions carried by event occurrences to a state value. It stores
the new state and outputs new event occurrences based on the output of the function.
The accumulator has a simple implementation using {\tt switch}.

The state stored is the table (implemented using the {\tt Data.Map} module in
the Haskell {\tt containers} library) and the output event occurrences carry
messages to be sent to the switch. The events are sourced internally to the
learning switch signal function (using the {\tt after} combinator in TimeFlies
and the {\tt repeatedly} function in Yampa) to delete expired rules, and
externally, by transforming messages from the switch to the controller into
accumulator functions. This implementation is shown in Figure~\ref{figure:openflow_implementation}.

\begin{figure}
\begin{code}
learn ::     SVEvent (SwitchHandle EthernetFrame,
                      TransactionID,
                      SCMessage EthernetFrame)
         :~> SVEvent (IO ())
learn = uncancelLeft >>>
        -- Timer to clean the table
        first (cleanRules tableTime) >>>
        -- Builds accumulator actions from input messages 
        second handleSCMessage >>>
        -- Bring together accumulator inputs     
        union >>> 
        -- The accumulator                       
        switchTable >>>
        -- Produce IO actions as event outputs                 
        sendCSMessage

cleanRules ::     Double
               -> SVEmpty
              :~> SVEvent TableAccumulator
cleanRules t = copy >>> -- two empty inputs
               -- Fire every t seconds
               second (every t cleanRulesAcc) >>> 
               -- capture the time each fires
               first time >>>
               capture >>>                        
               pureEventTransformer (uncurry ($))
  where 
    cleanRulesAcc :: Double -> TableAccumulator
    cleanRulesAcc nowT table = cleanTable t nowT table

-- | Produce an event with the given value at the given time interval
every :: Double -> a -> SVEmpty :~> SVEvent a
every t x = let e = switch (after t x >>>
                            copy >>>
                            second (pureEventTransformer $ const e))
            in e
\end{code}
\hrule
\caption{Partial implementation of the OpenFlow switch using TimeFlies}.
\label{figure:openflow_implementation}
\end{figure}

@@

\begin{figure}
\includegraphics{graph}
\hrule
\caption{Comparison of timeflies vs. Yampa implementing an OpenFlow learning switch controller.}
\label{fig:timeflies-yampa-comparison}
\end{figure}

### 2013-01-11 Docs/Thesis/introduction.tex
Take care of revisions from mtf

Functional reactive programming (FRP) is a promising class of abstractions for
writing reactive programs in functional languages. The FRP abstractions are
behaviors or signals, which are functions of continuous time, and events, which
are sequences of time-value pairs. These abstractions translate desirable
properties of the underlying functional languages, such as higher-order
functions and referential transparency, to reactive constructs, generally
without modifications to the underlying language. This permits the compositional
construction of reactive programs in purely functional languages.

@@

The first challenge was addressed by representing FRP programs, not as
compositions of signals and events, but as compositions of transformers of
behaviors and events called {\em signal functions}. By not permitting direct
access to behaviors or events, but representing them implicitly, signal
functions prohibit accessing past or future time values, only allowing access to
values in the closure of the signal function and current input values. Signal
function FRP programs are written by composing signal functions, and are
evaluated using a function which provides input to the signal function and acts
on the output values. This evaluation function is generally elided in existing
literature on signal functions, but we provide a documented and improved
interface for evaluating signal functions. A further advantage of signal
function FRP programs is that they are more readily composed, since additional
signal functions may be composed on the input or output side of a signal
function, rather than simply the output side as in classic (non-signal-function)
FRP systems.

>>>>

Functional reactive programming (FRP) is a promising class of abstractions for
writing interactive programs in functional languages. The FRP abstractions are
behaviors or signals, which are functions of continuous time, and events, which
are sequences of time-value pairs. These abstractions translate desirable
properties of the underlying functional languages, such as higher-order
functions and referential transparency, to reactive constructs, generally
without modifications to the underlying language. This permits the compositional
construction of reactive programs in purely functional languages.

@@

The first challenge was addressed by representing FRP programs, not as
compositions of signals and events, but as compositions of transformers of
behaviors and events called {\em signal functions}. By not permitting direct
access to behaviors or events, but representing them implicitly, signal
functions prohibit accessing past or future time values, only allowing access to
values in the closure of the signal function and current input values. Signal
function FRP programs are written by composing signal functions, and are
evaluated using a function which provides input to the signal function and acts
on the output values. This evaluation function is generally elided in existing
literature on signal functions, but we provide a documented and improved
interface for evaluating signal functions. A further advantage of
signal-function FRP programs is that they are more readily composed,
since additional signal functions may be composed on the input or output side of
a signal function, rather than simply the output side as in classic
(non-signal-function) FRP systems.

### 2013-01-11 Docs/Thesis/system_design_and_interface.tex
Take care of revisions from mtf

In previous systems such as Yampa, events were understood (and implemented) as option-valued
signals. This approach is undesirable for several reasons. The most pressing reason is
that it prohibits push-based evaluation of events, because events are embedded in the
samples of a signal and must be searched for when sampling.

@@

Only those combinators which modify or combine signal functions
({\tt (>>>)}, {\tt first}, {\tt second}) are reactive (may replace themselves in
response to events, and then only if they inherit their reactivity from the
signal function(s) which are inputs to combinator. The rest do not
react to or modify the input in any way, except to re-arrange it, copy it, or
discard it altogether.

@@

There are other formulations of a reactive combinator found in FRP literature
and libraries, which may be implemented using the one supplied. These are shown
in Figure~\ref{figure:alternate_switching_combinators} and may be provided in a
future version of the TimeFlies library.

@@

-- Supplied version in terms of alternate version
switch ::    (svIn :~> (svOut :^: SVEvent (svIn :~> svOut)))
          -> svIn :~> svOut
switch sf = switch_gen sf id

@@

The evaluation interface provides a modified state monad which holds a signal
function, together with some additional information, as its state (shown in Figure~\ref{figure:evaluation_state}.
Rather than monadic instructions to put and get the state, the monad provides instructions
to trigger an input event, update an input signal, and trigger sampling of
signals in the signal function. Additional state includes the current set of
modifications to the input signals (since the last sample) and a set of
handlers which actuate effects based on output events or changes to the output
signal.

>>>>

In previous systems such as Yampa, events were understood (and implemented) as
option-valued signals.\footnote{{\tt type Event a = Signal (Maybe a)}} This
approach is undesirable for several reasons. The most pressing reason is that it
prohibits push-based evaluation of events, because events are embedded in the
samples of a signal and must be searched for when sampling.

@@

Only those combinators which modify or combine signal functions
({\tt (>>>)}, {\tt first}, {\tt second}) are reactive (may replace themselves in
response to events), and then only if they inherit their reactivity from the
signal function(s) which are inputs to combinator. The rest do not
react to or modify the input in any way, except to re-arrange it, copy it, or
discard it altogether.

@@

There are other formulations of a reactive combinator found in FRP literature
and libraries, which may be implemented using the one supplied. Some of these
are shown in Figure~\ref{figure:alternate_switching_combinators} and may be
provided in a future version of the TimeFlies library.

@@

@@

The goal of the evaluation interface is to provide a means to evaluate
declaratively specified signal functions in the context of a monadic
input/output system, such as Haskell's IO monad. The means providing a simple
monadic interface for taking actions on a signal function in response to
inputs, and for actuating the outputs of a signal function.

The evaluation interface provides a modified state monad which holds a signal
function, together with some additional information, as its state (shown in
Figure~\ref{figure:evaluation_state}).
Rather than monadic instructions to put and get the state, the monad provides instructions
to trigger an input event, update an input signal, and trigger sampling of
signals in the signal function. Additional state includes the current set of
modifications to the input signals (since the last sample) and a set of
handlers which actuate effects based on output events or changes to the output
signal.
